Embedding Intuitionistic into Classical Logic

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
some typos

Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g. by assuming non-existence and deriving a contradiction.
The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in famous Curry-Howard correspondence~\cite{howard1980formulae}.

The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmarks problems (see e.g. the
ILTP library website~\cite{iltp}).
However, progress in automated theorem proving for intuitionistic logic has been slow whereas reasoners for classical logic have made tremendous progress, see e.g. the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g. TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate. Furthermore all the most popular first-order utilize classical logic.
This difference can be partially explained by fundamental differences between the logics.
First of all, determining intuitionistic validity is computationally harder, i.e. in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e. to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e. directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization. This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
Finally we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e. SAT-solvers.

With this work we want to open a path for leveraging the progress in classical theorem proving for intuitionistic logic.
To this end we give an embedding from intuitionistic logic into classical logic. That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case. In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid. Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent. This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e. if they both occur in $\varphi$ we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$. In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
%
In addition to the translation of formulas we give an effective translation of counter-models, i.e. for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
We note that the existence of counter-models is a key notion, which forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.

@@

\textbf{Related Work.}
Several approaches to obtain constructive content from classical proofs have been studied in the literature.
An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
Other very restricted subsets of classical logic have been identified under which this entailment holds, e.g. in~\cite{schwichtenberg}, however this approach is not applicable in general.
Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
In this case we necessarily lose realizability at least of $\exists$ and $\vee$ in general.

@@

We shall now give an overview of our main results and highlight the key arguments. Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. We start with the propositional case. The key arguments are the same as in the first-order case, while it is technically simpler and less cluttered.

@@

\begin{theorem}
	Let $\mathcal S$ be as above. There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable. Fixing $N\in\mathbb N$ deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
\end{theorem}

@@

\begin{theorem}
\label{thm:reduction-first-order-short}
There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. The size of $\varphi^\#$ is in linear in the size of $|\varphi|$, however for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$ and furthermore $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols. Furthermore there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
\end{theorem}
A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{fullFOtranslation}.

These results provide an effective framework for using classical provers to check intuitionistic validity. First experiments with the Vampire theorem prover show promise.

@@

\begin{definition}
A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
	\end{itemize}
We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
	\begin{itemize}
		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. A formula is valid if it is satisfied in eveery $\Sigma$-Kripke structure, it is valid for some Frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that Frame. We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
Analogous remarks with regard to validity and satisfiability hold as in the propositional case.

@@

\begin{remark}
	It is well known that Kripke Frames which are trees are complete with regard to intuitionistic logic. Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
\end{remark}

@@

	The other direction is a bit more involved. Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$. In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$. Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$. Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
	Now have
	\begin{itemize}
		\item $s^I = w_0$.
		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
		\item $f^I(m_1\dots m_n) =\begin{cases}
			f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
			w_0, & \text{otherwise.}
		\end{cases}$
		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
			\top, & \text{otherwise.}
		\end{cases}$
	\end{itemize}
	One easily verifies that these are well-defined. Let us now briefly check that this indeed defines a counter-model. First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is. Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified. All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$. Consider e.g. the case where $\psi\in\mathcal S$ is of the form
	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
	We have to show that
	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false. For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$. Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$. Note that the above is false in particular for $w\in M_0$ with $u\preceq w$. But this would imply
	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
\end{proof}

>>>>

Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g. by assuming non-existence and deriving a contradiction.
The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particul<ar motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in famous Curry-Howard correspondence~\cite{howard1980formulae}.

The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmarks problems (see e.g. the
ILTP library website~\cite{iltp}).
However, progress in automated theorem proving for intuitionistic logic has been slow, whereas reasoners for classical logic have made tremendous progress, see e.g. the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g. TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate. Furthermore all the most popular first-order utilize classical logic.
This difference can be partially explained by fundamental differences between the logics.
First of all, determining intuitionistic validity is computationally harder, i.e. in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e. to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e. directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization. This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e. SAT-solvers.

With this work we want to open a path for leveraging the progress in classical theorem proving for intuitionistic logic.
To this end, we give an embedding from intuitionistic logic into classical logic. That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case. In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid. Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent. This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e. if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$. In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
%
In addition to the translation of formulas we give an effective translation of counter-models, i.e. for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.

@@

\textbf{Related Work.}
Several approaches to obtain constructive content from classical proofs have been studied in the literature.
An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other settings have been studied~\cite{schwichtenberg}.
Another important direction is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}, which however does not permit in general the realizability of at least $\exists$ and $\vee$.

@@

We shall now give an overview of our main results and highlight the key arguments. Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. We start with the propositional case. The key arguments are the same as in the first-order case, while it is technically simpler and less cluttered.

@@

\begin{theorem}
	Let $\mathcal S$ be as above. There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable. For fixed $N\in\mathbb N$, deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
\end{theorem}

@@

\begin{theorem}
\label{thm:reduction-first-order-short}
There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols. Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
\end{theorem}
A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{fullFOtranslation}.

These results provide an effective framework for using classical provers to check intuitionistic validity. First experiments with the Vampire theorem prover~\cite{kovacs2013first} show promise.

@@

\begin{definition}
A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
	\end{itemize}
We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
	\begin{itemize}
		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that frame. We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
Analogous remarks with regard to validity and satisfiability hold as in the propositional case.

@@

\begin{remark}
	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic. Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
\end{remark}

@@

	The other direction is a bit more involved. Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$. In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$. Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$. Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
	Now have
	\begin{itemize}
		\item $s^I = w_0$.
		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
		\item $f^I(m_1\dots m_n) =\begin{cases}
			f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
			w_0, & \text{otherwise.}
		\end{cases}$
		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
			\top, & \text{otherwise.}
		\end{cases}$
	\end{itemize}
	One easily verifies that these are well-defined. Let us now briefly check that this indeed defines a counter-model. First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is. Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified. All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$. Consider e.g. the case where $\psi\in\mathcal S$ is of the form
	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
	We have to show that
	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false. For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$. Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$. Note that the above is false in particular for $w\in M_0$ with $u\preceq w$. But this would imply
	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
\end{proof}
